Parsing Postfix Log Files

### 2009-04-27 doc/logparser-results.tex
Corrections to chapter 6 from proofreading.

This chapter evaluates \parsername{} on two criteria: efficiency, and
coverage of  Postfix log files.  Efficiency is important because
\parsername{} is intended to be used in a production environment, and must
be capable of parsing log files generated by a high volume mail server in a
reasonable period of time.  Full coverage of Postfix log files is important
because the data gathered by the parser must be accurate and complete for
it to be useful and reliable.  The two goals are mutually incompatible:
every extra measure implemented to improve accuracy slows down parsing.
Some of the complications described in \sectionref{complications} occur
fewer than ten times when parsing the \numberOFlogFILES{} log files used to
evaluate \parsername{} in this chapter, yet their solutions slow down
parsing of every log file.  However, accurate but slow parsing is
preferable to sloppy but quick parsing.

The performance evaluation begins by describing the characteristics of the
mail server that produced the log files used to evaluate the parser's
performance and coverage; the computer that the tests were run on is also
described.  The characteristics of the \numberOFlogFILES{} log files are
described next, with an explanation of why \parsername{} has better
performance when parsing the larger log files in the group.  The effect of
using different rule orderings is explored, and the use of optimal rule
ordering is compared to use of an oracle that allows the parser to use only
one rule when recognising each log line.  When \parsername{} is used by
other mail administrators, they will need to extend the ruleset to parse
their own log lines, so the next section addresses the question of how
parser performance is affected by an increase in the number of rules in the
ruleset.  The penultimate topic covered is the simple optimisation of
caching the results of compiling each rule's regex, and the huge effect it
has on parser efficiency.  The performance evaluation concludes by
examining where parsing time is spent: recognition of log lines or their
subsequent processing?

The second criterion the parser is evaluated on is its coverage of Postfix
log files.  This subject contains two sections: what proportion of log
lines are correctly recognised by the ruleset, and what proportion of mail
delivery attempts are correctly understood and reconstructed by the
actions?  The former is a requirement for the latter to be achieved, and
the latter is important because the data provided by \parsername{} must be
both complete and correct for it to be of use to others.

@@

Efficiency is an obvious concern when the parser routinely needs to parse
large log files.  The mail server that generated the \numberOFlogFILES{}
log files is a production mail server handling mail for a university
department; the benefit of using this mail server is that its log files
exhibit the idiosyncrasies and peculiarities that a mail server in the wild
must deal with, but the downside is that significantly altering the
configuration to accommodate this project is not an option.
\Graphref{Mails received per day} and \tableref{Number of mails received
per day: statistics} show that, as expected, it accepts far more mails on
weekdays than at weekends.  Note that mail generated on the server does not
contribute to the figures: in particular, the vast amount of mail resulting
from the mail loops noticeable in later graphs is not reflected in these
figures.

@@

When generating the timing data used in this section, \numberOFlogFILES{}
log files were used, each containing one day's log lines, totalling three
months of contiguous log files,
\input{build/include-total-size-of-93-log-files}, \numberOFlogLINEShuman{}
log lines.  Each log file was parsed 10 times, and the mean parsing time
calculated.  The computer used for test runs was a Dell Optiplex 745, shown
in \tableref{Details of the computer used to generate statistics}; it was
dedicated to the task of gathering statistics from test runs, and did not
run any other programs while test runs were in progress.  Saving results to
the database was disabled for the test runs, because that dominates the run
time of the parser, and the tests are aimed at measuring the speed of
\parsername{} rather than the speed of the database and the disks it is
stored on.  Parsing all \numberOFlogFILES{} log files in one run without
saving results to the database took
\input{build/include-full-run-duration}, with mean throughput of
\input{build/include-full-run-throughput}.  In contrast, when saving
results to the database, parsing all \numberOFlogFILES{} log files took
\input{build/include-insert-results-duration}, with mean throughput of
\input{build/include-insert-results-throughput} --- parsing takes up only
\input{build/include-skip-inserting-as-percentage-of-inserting} of the
execution time when saving results to the database.

@@

The \numberOFlogFILES{} log files used when generating statistics have
fairly consistent sizes and contents, except for two groups of log files:
22 \& 62--68.  Median log file size is
\input{build/include-median-file-size}, containing
\input{build/include-median-number-of-log-lines} log lines, with a median
of \input{build/include-median-mails-accepted-over-smtp} mails accepted
each weekday; large scale mail servers would generate much larger log
files.  \Graphref{parsing time vs log file size vs number of log lines
graph} shows the parsing time in seconds, log file size in MB, and number
of log lines in tens of thousands, for each of the \numberOFlogFILES{} log
files.  Parsing time is plotted against number of log lines in
\graphref{Parsing time plotted against number of log lines}, and against
log file size in \graphref{Parsing time plotted against log file size};
the points on both graphs form almost perfectly straight lines.

\Graphref{parsing time vs log file size vs number of log lines
factor} plots both the ratio of log file size vs parsing time, and the
ratio of number of log lines vs parsing time (higher is more efficient in
both cases); \tableref{parsing time vs log file size vs number of log lines
factor table} shows the same ratios for different groups of log files.  The
ratios are quite tightly banded, except they increase (i.e.\ improve) for
log files 22 \& 62--68, despite their larger than usual size.  Both groups
of log files are much larger than usual because of mail loops caused by
users who set up mail forwarding incorrectly, resulting in a very different
distribution of log lines: normally most log lines are generated by mail
delivery attempts from other hosts, but when the mail loops occurred most
of the log lines resulted from failed delivery of mail generated on the
server itself.  The Postfix components that generated most of the log lines
during the mail loops have fewer associated rules than the Postfix
components whose log lines normally make up the bulk of each log file, so
the mean number of rules tried per log line is lower for these log files,
as is the mean parsing time per log line.  \Tableref{Number of rules per
Postfix component} shows the number of rules for each Postfix component;
\graphref{Mean number of rules used per log line} shows the drop in the
mean number of rules tried per log line for log files containing a mail
loop, and \graphref{Mean number of rules used per log line for each Postfix
component} shows the mean number of rules tried per log line for each
Postfix component for all log files.

@@

    \item [Shuffle] This ordering is intended to represent a randomly
        ordered rule set.  The rules will be shuffled once before use and
        will retain that ordering until the parser exits.  Note that the
        ordering will change every time the parser is executed, so 10
        different rule orderings will be generated for each log file in the
        test run.

@@

The parsing time of the three orderings are plotted against log file size
in \graphref{Parsing time plotted against log file size for optimal,
shuffled, and reverse orderings}; \graphref{Parsing time of optimal and
reverse orderings relative to shuffled ordering graph} shows the parsing
times of optimal and reverse orderings relative to shuffled ordering, and
mean relative parsing times for different groupings of log files are given
in \tableref{Parsing time of optimal and reverse orderings relative to
shuffled ordering table}.  This optimisation provides a mean reduction in
parsing time of
\input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
with normal log files,
\input{build/include-optimal-ordering-parsing-time-reduction-logs-22-62-68}\%
when a mail loop occurs and the distribution of log lines is unusual.
\Tableref{Parsing time of optimal and reverse orderings relative to
shuffled ordering table} shows that differences in rule ordering have less
effect on parsing time when parsing log files 22 \& 62--68, because of the
different distribution of log lines in those log files.  A careful
examination of \graphref{Parsing time of optimal and reverse orderings
relative to shuffled ordering graph} shows that, for the first log file
only, optimal and reverse orderings perform identically: this is because
the hits field of each rule is zero for the first log file, so optimal and
reverse orderings produce identical rule orderings.  For the first log
file, shuffled ordering is the most efficient of the three, but that is
accidental and cannot be relied upon.

@@

Optimal rule ordering, as described in \sectionref{rule ordering for
efficiency}, is the best rule ordering it is possible to achieve without
having an oracle that magically divines which rule should be used to
recognise each log line.  Such an oracle would give perfect performance,
because only one rule would need to be used to recognise each log line.
\parsername{} can save a list showing which rule recognised each log line,
and use that list to simulate an oracle and improve parsing speed the
\textit{second\/} time a log file is parsed.  This does not provide a
practical benefit, but it does provide a means to evaluate the performance
of optimal rule ordering in comparison to an oracle.

\Graphref{Parsing time of oracle and optimal ordering relative to shuffled
ordering graph} shows how the oracle and optimal ordering perform relative
to shuffled ordering, with mean and standard deviation in \tableref{Parsing
time of oracle and optimal ordering relative to shuffled ordering table}.
As expected, the oracle is more efficient than optimal ordering, but not by
much.  \Graphref{Percentage increase in parsing time when using optimal
ordering instead of oracle graph} shows the percentage increase in parsing
time when using optimal ordering instead of the oracle, with mean and
standard deviation in \tableref{Percentage increase in parsing time when
using optimal ordering instead of oracle table}.

@@

Optimal ordering proves to be quite efficient: \tableref{Percentage
increase in parsing time when using optimal ordering instead of oracle
table} shows that optimal ordering is less than
\input{build/include-perfect-best-vs-optimal-mean} slower than parsing
using an oracle that magically divines the correct rule to use for each log
line.

XXX TAKE UP SOME SPACE\@.

XXX TAKE UP SOME SPACE\@.

XXX TAKE UP SOME SPACE\@.

\showgraph{build/graph-perfect-best-and-optimal-vs-shuffled}{Parsing time
of oracle and optimal ordering relative to shuffled ordering}{Parsing time
of oracle and optimal ordering relative to shuffled ordering graph}

\showtable{build/include-perfect-best-and-optimal-vs-shuffle}{Parsing time
of oracle and optimal ordering relative to shuffled ordering}{Parsing time
of oracle and optimal ordering relative to shuffled ordering table}

\showgraph{build/graph-perfect-best-vs-optimal}{Percentage increase in
parsing time when using optimal ordering instead of oracle}{Percentage
increase in parsing time when using optimal ordering instead of oracle
graph}

\showtable{build/include-perfect-best-vs-optimal-stddev}{Percentage
increase in parsing time when using optimal ordering instead of
oracle}{Percentage increase in parsing time when using optimal ordering
instead of oracle table}

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important for this architecture because it is
anticipated that the typical parser will have a large ruleset.  The full
\parsername{} ruleset has \numberOFrules{} rules, whereas the minimum
ruleset required to parse the \numberOFlogFILES{} log files has
\numberOFrulesMINIMUM{} rules, \numberOFrulesMINIMUMpercentage{} of the
full ruleset.  The full ruleset is larger because \parsername{} is tested
with \numberOFlogFILESall{} log files; testing with more log files
increases the chance of finding bugs in the parser or new complications to
be overcome.  The \numberOFlogFILES{} log files were each parsed 10 times
using the minimum ruleset, and the mean parsing times compared to those
generated using the full ruleset: the percentage parsing time increase when
using the full ruleset instead of the minimal ruleset for optimal,
shuffled, and reverse orderings is shown in \graphref{Percentage parsing
time increase when using the maximum ruleset instead of the minimum
ruleset}, with mean and standard deviation in \tableref{Percentage parsing
time increase when using the maximum ruleset instead of the minimum ruleset
table}.  Note that in both \graphref{Percentage parsing time increase when
using the maximum ruleset instead of the minimum ruleset} and
\tableref{Percentage parsing time increase when using the maximum ruleset
instead of the minimum ruleset table}, for each ordering the parsing time
using the maximum ruleset is compared to the ordering's parsing time using
the minimum ruleset: the three orderings are not compared to a common
baseline.

Clearly the increased number of rules causes a noticeable performance
decrease with reverse ordering, and a lesser decrease with shuffled
ordering, whereas optimal ordering shows scant change.  Log files 22 \&
62--68 show much smaller increases in parsing time than other log files do,
because most of the log lines in those log files are produced by Postfix
components with few rules, so removing unnecessary rules has little effect
on the total number of rules used; \tableref{Number of rules per Postfix
component in the maximum and minimum rulesets} shows the number of rules
per Postfix component for each ruleset.  Once again, \graphref{Percentage
parsing time increase when using the maximum ruleset instead of the minimum
ruleset} shows that optimal and reverse orderings have identical
performance for the first log file, because the hits field of every rule
starts at zero.

@@

\Graphref{Percentage increase in parsing time when not caching compiled
regexes graph} shows the effect that not caching compiled regexes has on
parser performance, with mean and standard deviation in
\tableref{Percentage increase in parsing time when not caching compiled
regexes table}.  For typical log files, the mean increase in parsing time
when not caching compiled regexes is
\input{build/include-cached-regexes-vs-discarded-regexes-mean}.  Caching
compiled regexes is probably the single most effective optimisation
possible in \parsername{}, and was quite simple to implement: the framework
compiles each rule's regex when the ruleset is loaded, and uses the
compiled regex instead of the source regex when recognising log lines.  As
seen previously, log files 22 \& 62--68 do not suffer such a large increase
in parsing time when the optimisation is disabled; this is because, on
average, fewer rules are used so fewer regexes are compiled per log line
for those log files.  The increase in parsing time when parsing the first
log file is much greater than for the other log files; again, this is
because every rule's hits field starts at zero, so optimal ordering is less
efficient than usual, the mean number of rules used for each log line is
higher than usual, and more regexes will need to be compiled when
recognising each log line.

@@

@@

\subsection{Where Is The Time Spent: Recognising Or Processing Log Lines?}

@@

The optimisations described in this chapter have addresses the process of
recognising log lines, but have not optimised actions at all.  Optimisation
efforts have concentrated on recognition of log lines for two reasons:

@@

    \item Even with the optimisations described in this chapter enabled,
        recognising log lines still accounts for two thirds of the parser's
        execution time.  \Graphref{Percentage of parsing time spent
        recognising log lines graph} shows the percentage of parsing time
        spent recognising log lines for each of the \numberOFlogFILES{} log
        files, with mean and standard deviation shown in
        \tableref{Percentage of parsing time spent recognising log lines
        table}; for normal log files,
        \input{build/include-percentage-time-spent-recognising-log-lines-mean}
        of parsing time is spent recognising log lines.  To measure how
        long recognition of log lines takes, a full test run was performed
        as previously described in \sectionref{parser efficiency}, but with
        one difference: actions were not invoked when a log line was
        recognised, so the parsing time did not include the time normally
        taken by invocation of actions; the mean parsing times for each log
        file were subtracted from the mean parsing time from a normal test
        run to obtain the parsing time taken by actions.

@@

Individual actions or the framework could and have been optimised, but
plenty of existing literature is available on the topic of optimising
programs, so the subject is not dealt with here.

@@

    \item Try each rule for that Postfix component, then generic rules,
        until a recognising rule is found; if the log line is not
        recognised, issue a warning and move on to the next log line.

@@

Each Postfix component whose log lines are of interest must have at least
one rule, or all of its log lines will be silently skipped; in the extreme
case of an empty ruleset the parser would skip every log line.
\parsername{} skips log lines logged by programs that do not have any
associated rules, because there may be any number of log lines from other
programs in the log file, and some Postfix components do not produce any
log lines of interest.  \parsername{} does not parse log lines from
non-Postfix programs, e.g.\ Amavisd-new or SpamAssassin; it could easily be
extended to do so, if a method could be developed to correctly associate
such log lines with existing state table entries.  To correctly recognise
all log lines, there must be a rule to recognise each log line variant
produced by each Postfix component; if a log line is not recognised the
parser will issue a warning, to inform the user that they need to extend
their ruleset.  Each rule's regex should be as specific and precise as
possible, to ensure accurate parsing: a rule with a regex that matches zero
or more of any character will recognise every log line, but not in a
meaningful way; most log lines contain fixed strings, so this is not a
problem in practice.

Full coverage of log lines can be achieved without undue effort, yet once
achieved it requires maintenance.  Maintaining full coverage takes effort because
the set of log line variants changes over time, e.g.\ administrators add
restrictions with custom messages, \acronym{DNSBL} messages change, or
major releases of Postfix change log lines (usually by adding more
information).  Warnings are issued for any log lines that are not
recognised; no warnings are issued for unrecognised log lines while parsing
the \numberOFlogFILES{} log files, so it can be safely concluded that zero
false negatives arise.  False positives are harder to quantify, short of
examining each of the 60,721,709 log lines and checking that the correct
rule recognised it.  However, a random sample of 6039 log lines was parsed,
and the results manually verified by inspection to ensure that the correct
rule recognised each log line.  The sample was generated by running the
command: \texttt{perl -n -e \singlequote{}print if (rand 1 <
0.0001)\singlequote{} LOG\_FILES} to randomly extract roughly one log line
in every 10,000 (it actually extracted 0.00994\% instead of 0.01\%).  Each
log line was examined and the correct rule identified from the
\numberOFrules{} rules in the database; the correct rule was then compared
to the rule that recognised the log line when parsing.  The sample results
contained zero false positives, and this check has been automated to ensure
continued accuracy.  Based on these results, and how precise each rule's
regex is, the author is confident that zero false positives occur when
parsing the \numberOFlogFILES{} log files.

@@

    \item 9 mails whose initial log lines were not present in the log
        files.  These mails did not produce warnings because they resemble
        child mails waiting to be tracked with a parent; see
        \sectionref{Re-injected mails} for details.

    \item 7 mails that had yet to be delivered because of repeated
        failures.

@@

    \item \parsername{} performs many checks to detect known problems
        occurring, e.g.\ if a mail is missing required data.  No such
        warnings are produced during the test runs other than the five
        described above.

    \item Queueids and \glspl{pid} naturally identify log lines belonging
        to one mail or connection respectively; it is extremely unlikely
        that a log line would be associated with the wrong connection.

>>>>

This chapter evaluates \parsername{} on two criteria: efficiency, and
coverage of  Postfix log files.  Efficiency is important because
\parsername{} is intended to be used in a production environment, and must
be capable of parsing log files generated by a high volume mail server in a
reasonable period of time.  Full coverage of Postfix log files is important
because the data gathered by the parser must be accurate and complete for
it to be useful and reliable.  Progress in achieving one goal usually comes
at the expense of the other: every extra measure implemented to improve
accuracy slows down parsing.  Some of the complications described in
\sectionref{complications} occur fewer than ten times when parsing the
\numberOFlogFILES{} log files used to evaluate \parsername{} in this
chapter, yet their solutions slow down parsing of every log file.  However,
those solutions are retained in \parsername{}, because accurate but slow
parsing is preferable to sloppy but quick parsing.

The performance evaluation begins by describing the characteristics of the
mail server that produced the \numberOFlogFILES{} log files used to
evaluate the parser's performance and coverage; the computer that the tests
were run on is also described.  The characteristics of the
\numberOFlogFILES{} log files are described next, with an explanation of
why \parsername{} has better performance when parsing the larger log files
in the group.  The effect of using different rule orderings is explored,
and optimal rule ordering is compared to an oracle that allows the parser
to use only one rule when recognising each log line.  When \parsername{} is
used by other mail administrators, they will need to extend the ruleset to
parse their own log lines, so the next section addresses the question of
how parser performance is affected by an increase in the number of rules in
the ruleset.  The penultimate topic is the simple optimisation of caching
the results of compiling each rule's regex, and the huge effect it has on
parser efficiency.  The performance evaluation concludes by examining where
parsing time is spent: recognition of log lines or their subsequent
processing by actions?

The second criterion the parser is evaluated on is its coverage of Postfix
log files.  There are two types of coverage to be evaluated: what
proportion of log lines are correctly recognised by the ruleset, and what
proportion of mail delivery attempts are correctly understood and
reconstructed by the actions?  The former is a requirement for the latter
to be achieved, and the latter is important because the data provided by
\parsername{} must be both complete and correct for it to be of use to
others.

@@

Efficiency is an obvious concern when the parser routinely needs to parse
large log files.  The mail server that generated the \numberOFlogFILES{}
log files used in this chapter is a production mail server handling mail
for a university department; the benefit of using this mail server is that
its log files exhibit the idiosyncrasies and peculiarities that a mail
server in the wild must deal with, but the downside is that significantly
altering the configuration to accommodate this project is not an option.
\Graphref{Mails received per day} and \tableref{Number of mails received
per day: statistics} show that, as expected, it accepts far more mails on
weekdays than at weekends.  Note that mail generated on the server (bounce
notifications, mail re-injected for forwarding, mail sent from mailing
lists) does not contribute to these figures: in particular, the vast amount
of mail resulting from the mail loops noticeable in later graphs is not
reflected in these figures.

@@

When generating the timing data used in this section, \numberOFlogFILES{}
log files were used, each containing one day's log lines; the three months
of contiguous log files contain \numberOFlogLINEShuman{} log lines and
total \input{build/include-total-size-of-93-log-files}.  Each log file was
parsed 10 times, and the mean parsing time calculated for each log file.
The computer used for test runs was a Dell Optiplex 745, shown in
\tableref{Details of the computer used to generate statistics}; it was
dedicated to the task of gathering statistics from test runs, and did not
run any other programs while test runs were in progress.  Saving results to
the database was disabled for the test runs, because that dominates the run
time of the parser, and the tests are aimed at measuring the speed of
\parsername{} rather than the speed of the database and the disks it is
stored on.  Parsing all \numberOFlogFILES{} log files in one run without
saving results to the database took
\input{build/include-full-run-duration}, with mean throughput of
\input{build/include-full-run-throughput}.  In contrast, when saving
results to the database, parsing all \numberOFlogFILES{} log files took
\input{build/include-insert-results-duration}, with mean throughput of
\input{build/include-insert-results-throughput} --- parsing takes up only
\input{build/include-skip-inserting-as-percentage-of-inserting} of
execution time when saving results to the database.

@@

The \numberOFlogFILES{} log files used when generating statistics have
fairly consistent sizes and contents, except for two groups of log files:
22 \& 62--68.  Median log file size is
\input{build/include-median-file-size}, containing
\input{build/include-median-number-of-log-lines} log lines, with a median
of \input{build/include-median-mails-accepted-over-smtp} mails accepted
each weekday; large scale mail servers would accept many more mails and
consequently generate much larger log files.  \Graphref{parsing time vs log
file size vs number of log lines graph} shows the parsing time in seconds,
log file size in MB, and number of log lines in tens of thousands, for each
of the \numberOFlogFILES{} log files.  Parsing time is plotted against
number of log lines in \graphref{Parsing time plotted against number of log
lines}, and against log file size in \graphref{Parsing time plotted against
log file size}; the points on both graphs appear to form straight lines,
but they actually curve downwards slightly as they move to the right.

\Graphref{parsing time vs log file size vs number of log lines factor}
plots both the ratio of log file size to parsing time, and the ratio of
number of log lines to parsing time (higher is more efficient in both
cases); \tableref{parsing time vs log file size vs number of log lines
factor table} shows the same ratios for different groups of log files.  The
ratios are quite tightly banded, except they increase (i.e.\ improve) for
log files 22 \& 62--68, despite their larger than usual size.  Both groups
of log files are much larger than usual because of mail loops caused by
users who set up mail forwarding incorrectly, resulting in a very different
distribution of log lines: normally most log lines are generated by mail
delivery attempts from other hosts, but when the mail loops occurred most
of the log lines resulted from failed delivery of mail re-injected for
forwarding.  The Postfix components that generated most of the log lines
during the mail loops have fewer associated rules than the Postfix
components whose log lines normally make up the bulk of each log file, so
the mean number of rules used per log line is lower for these log files, as
is the mean parsing time per log line.  \Tableref{Number of rules per
Postfix component} shows the number of rules for each Postfix component;
\graphref{Mean number of rules used per log line} shows the drop in the
mean number of rules used per log line for log files containing a mail
loop, and \graphref{Mean number of rules used per log line for each Postfix
component} shows the mean number of rules used per log line for each
Postfix component for all log files.

@@

    \item [Shuffle] This ordering is intended to represent a randomly
        ordered ruleset.  The rules will be shuffled once before use and
        will retain that ordering until the parser exits.  Note that the
        ordering will change every time the parser is executed, so 10
        different rule orderings will be generated for each log file in the
        test run.

@@

The parsing times of the three orderings are plotted against log file size
in \graphref{Parsing time plotted against log file size for optimal,
shuffled, and reverse orderings}; \graphref{Parsing time of optimal and
reverse orderings relative to shuffled ordering graph} shows the parsing
times of optimal and reverse orderings relative to shuffled ordering, and
mean relative parsing times for different groupings of log files are given
in \tableref{Parsing time of optimal and reverse orderings relative to
shuffled ordering table}.  This optimisation provides a mean reduction in
parsing time of
\input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
with normal log files.  \Tableref{Parsing time of optimal and reverse
orderings relative to shuffled ordering table} shows that differences in
rule ordering have less effect on parsing time when parsing log files 22 \&
62--68, because of the different distribution of log lines in those log
files.  A careful examination of \graphref{Parsing time of optimal and
reverse orderings relative to shuffled ordering graph} shows that, for the
first log file only, optimal and reverse orderings perform identically:
this is because the hits field of each rule is zero for the first log file,
so optimal and reverse orderings produce identical rule orderings.  For the
first log file, shuffled ordering is the most efficient of the three, but
that is accidental and cannot be relied upon.

@@

Optimal rule ordering, described in \sectionref{rule ordering for
efficiency}, is the best rule ordering it is possible to achieve without
having an oracle that magically divines which rule should be used to
recognise each log line.  Such an oracle would give perfect performance,
because only one rule would need to be used to recognise each log line.
\parsername{} can save a list showing which rule recognised each log line,
and use that list to simulate an oracle and improve parsing speed the
\textit{second\/} time a log file is parsed.  This does not provide a
practical benefit, but it does provide a means to evaluate the performance
of optimal rule ordering in comparison to an oracle.

\Graphref{Parsing time of the oracle and optimal ordering relative to
shuffled ordering graph} shows how the oracle and optimal ordering perform
relative to shuffled ordering, with mean and standard deviation in
\tableref{Parsing time of the oracle and optimal ordering relative to
shuffled ordering table}; \graphref{Parsing time plotted against log file
size for the oracle, optimal ordering, and shuffled ordering} shows parsing
time plotted against log file size for the oracle, optimal ordering, and
shuffled ordering.  As expected, the oracle is more efficient than optimal
ordering, but not by much, particularly when parsing the larger log files.
\Graphref{Percentage increase in parsing time when using optimal ordering
instead of the oracle graph} shows the percentage increase in parsing time
when using optimal ordering instead of the oracle, with mean and standard
deviation in \tableref{Percentage increase in parsing time when using
optimal ordering instead of the oracle table}.

@@

Optimal ordering proves to be quite efficient: \tableref{Percentage
increase in parsing time when using optimal ordering instead of the oracle
table} shows that optimal ordering is less than
\input{build/include-perfect-best-vs-optimal-mean} slower than parsing
using an oracle that magically divines the correct rule to use for each log
line.

\showgraph{build/graph-perfect-best-and-optimal-vs-shuffled}{Parsing time
of the oracle and optimal ordering relative to shuffled ordering}{Parsing
time of the oracle and optimal ordering relative to shuffled ordering
graph}

\showtable{build/include-perfect-best-and-optimal-vs-shuffle}{Parsing time
of the oracle and optimal ordering relative to shuffled ordering}{Parsing
time of the oracle and optimal ordering relative to shuffled ordering
table}

\showgraph{build/graph-file-size-oracle-optimal-shuffle}{Parsing time
plotted against log file size for the oracle, optimal ordering, and
shuffled ordering}{Parsing time plotted against log file size for the
oracle, optimal ordering, and shuffled ordering}

\showgraph{build/graph-perfect-best-vs-optimal}{Percentage increase in
parsing time when using optimal ordering instead of the oracle}{Percentage
increase in parsing time when using optimal ordering instead of the oracle
graph}

\showtable{build/include-perfect-best-vs-optimal-stddev}{Percentage
increase in parsing time when using optimal ordering instead of the
oracle}{Percentage increase in parsing time when using optimal ordering
instead of the oracle table}

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important for this architecture because it is
anticipated that the typical parser will have a large ruleset.  The full
\parsername{} ruleset has \numberOFrules{} rules, whereas the minimum
ruleset required to parse the \numberOFlogFILES{} log files has
\numberOFrulesMINIMUM{} rules, \numberOFrulesMINIMUMpercentage{} of the
full ruleset.  The full ruleset is larger because \parsername{} is tested
with \numberOFlogFILESall{} log files (\numberOFlogFILESallYEARS{} of log
files); testing with more log files increases the chance of finding bugs in
the parser or new complications to be overcome.  The \numberOFlogFILES{}
log files were each parsed 10 times using the minimum ruleset, and the mean
parsing times compared to those generated using the full ruleset: the
percentage parsing time increase when using the full ruleset instead of the
minimal ruleset for optimal, shuffled, and reverse orderings is shown in
\graphref{Percentage parsing time increase when using the maximum ruleset
instead of the minimum ruleset}, with mean and standard deviation in
\tableref{Percentage parsing time increase when using the maximum ruleset
instead of the minimum ruleset table}.  For each ordering, the parsing time
using the maximum ruleset is compared to the parsing time using the minimum
ruleset: the three orderings are not compared to a common baseline.

Clearly, the increased number of rules causes a noticeable performance
decrease with reverse ordering, and a lesser decrease with shuffled
ordering, whereas optimal ordering shows scant change.  Log files 22 \&
62--68 show much smaller increases in parsing time than other log files do,
because most of the log lines in those log files are produced by Postfix
components with few rules, so removing unnecessary rules has little effect
on the total number of rules used; \tableref{Number of rules per Postfix
component in the maximum and minimum rulesets} shows the number of rules
per Postfix component for each ruleset.  Once again, \graphref{Percentage
parsing time increase when using the maximum ruleset instead of the minimum
ruleset} shows that optimal and reverse orderings have identical
performance for the first log file, because the hits field of every rule
starts at zero, and the two rule orderings are identical.

@@

A test run using optimal ordering was performed as described in
\sectionref{parser efficiency}, with one difference: regexes were not
compiled and cached when the ruleset was loaded from the database, so
Perl's regex engine had to compile each regex each time it was used when
trying to recognise a log line.  \Graphref{Percentage increase in parsing
time when not caching compiled regexes graph} shows the effect that not
caching compiled regexes has on parser performance, with mean and standard
deviation in \tableref{Percentage increase in parsing time when not caching
compiled regexes table}.  For typical log files, the mean increase in
parsing time when not caching compiled regexes is
\input{build/include-cached-regexes-vs-discarded-regexes-mean}; looking at
it from the opposite direction, caching compiled regexes reduces parsing
time by
\input{build/include-cached-regexes-vs-discarded-regexes-mean-reduction}.
Caching compiled regexes is probably the single most effective optimisation
possible in \parsername{}, and was quite simple to implement: the framework
compiles each rule's regex when the ruleset is loaded, and uses the
compiled regex instead of the source regex when recognising log lines.  As
seen previously, log files 22 \& 62--68 do not suffer such a large increase
in parsing time when the optimisation is disabled; this is because, on
average, fewer rules are used per log line, so fewer regexes are compiled
per log line for those log files.

The increase in parsing time when parsing the first log file is much
greater than for the other log files (see \graphref{Percentage increase in
parsing time when not caching compiled regexes graph}); again, this is
because every rule's hits field starts at zero, so optimal ordering is less
efficient than usual, the mean number of rules used for each log line is
higher than usual, and more regexes will need to be compiled when
recognising each log line.  Parsing time is plotted against log file size
for cached and discarded regexes in \graphref{Parsing time plotted against
log file size for cached regexes and discarded regexes}; the plot for
discarding compiled regexes is quite uneven compared to caching compiled
regexes --- some log files are particularly slow, whereas others are not as
badly affected, but these differences have not been investigated.

@@

\showgraph{build/graph-file-size-optimal-discarded}{Parsing time plotted
against log file size for cached regexes and discarded regexes}{Parsing
time plotted against log file size for cached regexes and discarded
regexes}

@@

\subsection{Where Is Parsing Time Spent: Recognising Or Processing Log
Lines?}

@@

The optimisations described in this chapter have addressed the process of
recognising log lines, but have not optimised actions at all.  Optimisation
efforts have concentrated on recognition of log lines for two reasons:

@@

    \item Even with the optimisations described in this chapter enabled,
        recognising log lines still accounts for two thirds of the parser's
        execution time.  \Graphref{Percentage of parsing time spent
        recognising log lines graph} shows the percentage of parsing time
        spent recognising log lines for each of the \numberOFlogFILES{} log
        files, with mean and standard deviation shown in
        \tableref{Percentage of parsing time spent recognising log lines
        table}; for normal log files,
        \input{build/include-percentage-time-spent-recognising-log-lines-mean}
        of parsing time is spent recognising log lines.  To measure how
        long recognition of log lines takes, a full test run was performed
        as previously described in \sectionref{parser efficiency}, but with
        one difference: actions were not invoked when a log line was
        recognised, so the parsing time did not include the time normally
        taken by invocation of actions; the mean parsing times for each log
        file were subtracted from the mean parsing times from a normal test
        run to obtain the parsing time taken by actions.  The parsing time
        exclude the time taken for framework initialisation, loading and
        saving state tables, and loading the ruleset: in ad far as
        possible, the times are just for recognising log lines or
        recognising log lines plus invoking actions.

@@

Individual actions and the framework could and have been optimised, but
plenty of existing literature is available on the topic of optimising
programs, so the subject is not dealt with here.

@@

    \item Try each rule that recognises log lines from that Postfix
        component, then any generic rules, until a recognising rule is
        found; if the log line is not recognised, issue a warning and move
        on to the next log line.

@@

Each Postfix component whose log lines are of interest must have at least
one rule that recognises its log lines, or all of its log lines will be
silently skipped; in the extreme case of an empty ruleset the parser would
skip every log line.  \parsername{} skips log lines from programs that do
not have any associated rules, because there may be any number of log lines
from other programs in the log file, and some Postfix components do not
produce any log lines of interest.  \parsername{} does not parse log lines
from non-Postfix programs, e.g.\ Amavisd-new or SpamAssassin; it could
easily be extended to do so, if a method could be developed to correctly
associate such log lines with existing state table entries.  To correctly
recognise all log lines, there must be a rule to recognise each log line
variant produced by each Postfix component; if a log line is not recognised
the parser will issue a warning, to inform the user that they need to
extend their ruleset.  Each rule's regex should be as specific and precise
as possible, to ensure accurate parsing: a rule with a regex that matches
zero or more of any character will recognise every log line, but not in a
meaningful way; most log lines contain fixed strings, so this is not a
problem in practice.

Full coverage of log lines can be achieved without undue effort, yet once
achieved it requires maintenance.  Maintaining full coverage is an ongoing
task because the set of log line variants changes over time, e.g.\
administrators add restrictions with custom messages, \acronym{DNSBL}
messages change, or major releases of Postfix change log lines (usually by
adding more information).  Warnings are issued for any log lines that are
not recognised; no warnings are issued for unrecognised log lines while
parsing the \numberOFlogFILES{} log files, so it can be safely concluded
that zero false negatives arise.  False positives are harder to quantify,
short of examining each of the 60,721,709 log lines and checking that the
correct rule recognised it.  However, a random sample of 6039 log lines was
parsed, and the results manually verified by inspection to ensure that the
correct rule recognised each log line.  The sample was generated by running
the command: \newline{} \tab{} \texttt{perl -n -e \singlequote{}print if
(rand 1 < 0.0001)\singlequote{} LOG\_FILES} \newline{} to randomly extract
roughly one log line in every 10,000 (it actually extracted 0.00994\%
instead of 0.010\%).  Each log line was examined and the correct rule
identified from the \numberOFrules{} rules in the database; the correct
rule was then compared to the rule that recognised the log line when
parsing.  The sample results contained zero false positives, and this check
has been automated to ensure continued accuracy.  Based on these results,
and how precise each rule's regex is, the author is confident that zero
false positives occur when parsing the \numberOFlogFILES{} log files.

@@

    \item 9 mails whose initial log lines were not present in the log
        files.  Some of those mails did not produce warnings because they
        resemble child mails waiting to be tracked with a parent; see
        \sectionref{Re-injected mails} for details.

    \item 7 mails that had yet to be delivered because of repeated
        delivery failures.

@@

    \item \parsername{} performs many checks to detect known problems,
        e.g.\ if a mail is missing required data.  No such warnings are
        produced during the test runs other than the five described above.

    \item Queueids and \glspl{pid} naturally identify log lines belonging
        to one mail or connection respectively; it is extremely unlikely
        that a log line would not be associated with the right connection.
